On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-04-15 MS.tex
Editorial changes Add Graphical Abstract Change color chart

\bibliographystyle{mee}
\mstype{Research Article}
\keywords{maximum height, photosynthetic rate, ontogenetic stage, growth strategy, seed mass, specific leaf area, wood density}

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 309/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate is seen as a central element of its ecological strategy, and as determined by its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of some prominent traits to influence growth rate.

@@

\item Synthesis: We show for some traits, how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from categorizing species as having slow or fast growth throughout their life, in favour of attributing growth trajectories.
\end{enumerate}

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth; and then ultimately replacing themselves with offspring. The idea that strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rates among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. During the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. This approach, using measurable traits, made broad comparisons possible across continents, latitudes, and thousands of species. Since then, functional traits have been seen as constructive tools to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of trait variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, recently a number of empirical studies indicate the correlation between traits and growth is not consistent. Consider SLA as an example. By definition, species with high SLA can deploy more leaf area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In seedlings growing under favourable conditions, SLA has repeatedly been found to be a strong predictor of relative growth rate, regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

@@

In this paper, we consider five prominent traits (specific leaf area = SLA, wood or stem tissue density, maximum leaf photosynthetic capacity per unit area = Aarea, maximum or asymptotic height, and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experimental study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

The rate at which \textbf{live biomass} $B$ is produced by a plant is given by
\begin{equation}\label{eq:eq0}
\underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}} = \underbrace{\alpha_{\rm y}}_{\text{yield}}  \big(\underbrace{A_{\rm l} p}_{\text{photosynthesis}} - \underbrace{\sum_{i = {\rm l,s,r}} M_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = \rm{l,s,r}} M_{i} k_{i}}_{\text{turnover}} \, ,
\end{equation}
where $\alpha_{\rm y}$ is the ratio of carbon fixed in mass per carbon assimilated, and $p$ is the assimilation rate of CO$_{2}$ per leaf area. Here, and in the following equations, $M_i$ denotes the \textbf{live mass on the plant}, $r_i$ the mass-specific \textbf{respiration rate}, and $k_i$ the \textbf{turnover rate} of different plant tissues; the subscripts $i = \rm{l,s,r,a,f}$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.)  Similarly $A_i$ denotes \textbf{areas}, of leaves ($i = l$) and of cross-sections of total stem and sapwood ($i =\rm{st,ss}$) respectively. Note that the live biomass $B = M_{\rm a} + M_{\rm f}$.  Applying the chain rule of differential calculus, the \textbf{growth rate in plant mass}, ($M_{\rm a} $), can be decomposed as
\begin{equation}\label{eq:eq1}
\underbrace{{\frac{\ud M_{\rm a} }{\ud t}}}_{\text{growth rate in mass}} = \underbrace{\frac{\ud M_{\rm a} }{\ud B}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}},
\end{equation}
where $\frac{\ud M_{\rm a} }{\ud B}$ is the fraction of net live biomass increase that goes to vegetative growth rather than to production of reproductive tissues.

The \textbf{growth rate in height} can likewise be decomposed as the product of several meaningful terms:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{\ud H}{\ud t}}_{\text{growth rate in height}} = \underbrace{\frac{\ud H}{\ud A_{\rm l}}}_{\text{architecture layout}} \times \underbrace{{\frac{\ud A_{\rm l}}{\ud M_{\rm a} }}}_{\text{leaf deployment per mass}} \times {\frac{\ud M_{\rm a} }{\ud B}} \times {\frac{\ud B}{\ud t}}
\end{equation}
where $\ud H / \ud A_{\rm l}$ is the growth in plant height
per unit growth in total leaf area. This term expresses the architectural
strategy of the plant. We assume it is unrelated to the traits considered in the current paper, in other words that it does not automatically change in response to changes in the traits. The term $\ud A_{\rm l} / \ud M_{\rm a} $
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures.  $\ud A_{\rm l} / \ud M_{\rm a} $
can itself be expressed as a sum of costs per unit leaf
area for constructing various support tissues,
\begin{equation}\label{eq:daldmt}
\frac{\ud A_{\rm l}}{\ud M_{\rm a} }
= \left(\frac{\ud M_{\rm l}}{\ud A_{\rm l}}
 +  \frac{\ud M_{\rm s}}{\ud A_{\rm l}} + \frac{\ud M_{\rm r}}{\ud A_{\rm l}}\right)^{-1}.
\end{equation}

Similarly \textbf{growth rate in basal area} can be expressed as the sum of growth in sapwood and heartwood, with the former term again decomposed as a product:
\begin{equation}\label{eq:eq3}
\begin{array}{ccccc}
\underbrace{{\frac{\ud A_{\rm st}}{\ud t}}}_{\text{basal area growth rate}} &=& \frac{\ud A_{\rm ss}}{\ud t} &+& \frac{\ud A_{\rm sh}}{\ud t} \\
&=& \underbrace{\frac{\ud A_{\rm ss}}{\ud A_{\rm l}}  \times {\frac{\ud A_{\rm l}}{\ud M_{\rm a} }} \times  {\frac{\ud M_{\rm a} }{\ud B}} \times {\frac{\ud B}{\ud t}}}_{\text{increase in total sapwood}} &+&  \underbrace{k_{\rm ss} \times \frac{\ud A_{\rm ss}}{\ud t}}_{\text{sapwood turnover}}.\\
\end{array}
\end{equation}
Here $\frac{\ud A_{\rm ss}}{\ud A_{\rm l}} $ is the  sapwood area per leaf area. The \textbf{growth rate in diameter} is then obtained via the geometric relationship between stem diameter and stem area,
\begin{equation}\label{eq:eq4}
\underbrace{\frac{\ud D}{\ud t}}_{\text{growth rate in diameter}} = \left(\pi \, A_{\rm st}\right)^{-0.5} \times {\frac{\ud A_{\rm st}}{\ud t}}.
\end{equation}

@@

The above equations are general, i.e. they do not depend on detailed assumptions about plant construction. To make qualitative predictions with respect to traits such as specific leaf area and wood density we require one additional assumption: that the mass of sapwood tissue ($M_{\rm ss}$) increases disproportionately with respect to the plant's leaf area ($A_{\rm l}$). This assumption is included in nearly every model of plant growth that has been developed, on the basis that as stems grow longer the total mass of sapwood associated with each leaf must be increasing \citep[e.g.][]{Falster:2011ii,King-1999,Enquist:2007ek}. It has the effect of changing the relative importance of different components within equations for $\frac{\ud B}{\ud t}$ (net live biomass production) and $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ (marginal cost of deploying additional leaf area) as a plant increases in size. Specifically, in eq. \ref{eq:eq0}, the costs of stem respiration and turnover increase faster than leaf photosynthesis, respiration or turnover, as the plant grows. Similarly, in eq. \ref{eq:daldmt}, the costs of constructing sapwood to support a given unit of leaf area, $\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$, increases faster than the cost of producing a unit of leaf area, $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ which is simply the inverse of specific leaf area. Whether plant growth rate increases with a specific trait, therefore depends on how the trait affects the different elements of \ref{eq:eq0}-\ref{eq:eq4} and how those effects change with size (Table \ref{tab:trade-offs}).

@@

\subsection*{H5: The relationship between Aarea and growth rate is expected to be positive across all plant sizes, particularly under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where Aarea is typically measured. This effect occurs by increasing  $\frac{\ud B}{\ud t} $ in eqs. (\ref{eq:eq1}-\ref{eq:eq3}). The positive effect of increasing Aarea (higher assimilation) may be offset by the costs (higher respiration), however, assuming both increase in proportion, and because assimilation is typically higher than respiration in high light, the net effect will be positive (Table \ref{tab:trade-offs}). It is uncertain how much this effect might translate to faster growth in low light.

@@

A big challenge when searching literature databases is to identify suitable studies from the vast corpus available. Initially we were interested to know whether data existed for a variety of traits; however the number of potential search terms made systemic searching for all these terms in Google Scholar impractical. A preliminary investigation of search results for a variety of traits on the Web of Science platform suggested that searching only on SLA and wood density -- two of the most widely measured traits -- would capture $\sim$80\% of search records that would be returned by searching for a wider range of traits (see Appendix \ref{app:literature-snapshot}). We therefore narrowed our searching to only cover these two main traits.

@@

The 19 traits recorded (Fig. \ref{fig:fig1}) capture different components of plant ecological strategies. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass and maximum height capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here wood density and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf versus other tissue types \citep{Lambers:1992bj}.

@@

We used categorical groupings by ontogenetic ``stage'' to test our hypotheses regarding the effects of traits on growth rates. The authors of the 103 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. Because these categories were established according to the expertise of the authors, we were concerned about the risk of assigning plants of similar size to different categories when compiling data across multiple studies and vegetation types. This risk turned out to be small: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages were relatively compatible with thresholds we defined ourselves (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in stem diameter (at 10 cm). Comparing these values to the bivariate relationships observed between variables (age, height, diameter) in the ``Biomass And Allometry Database'' \citep{Falster:2015}, demonstrated that the thresholds identified above for each stage transition also aligned relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries.

@@

\subsection*{Representation of ontogenetic stages among 19 traits}
In total, correlations had been reported for at least two different stages across 17 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits, correlations would be reported only for one growth stage. But this limitation applied only for NARmass and vessel density, where studies appeared confined to seedling and adult stages respectively. However, for a majority of traits the number of observations across stage was highly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from the same stage (Fig. \ref{fig:fig1}).

@@

For wood density and seed mass, the available results are consistent with our theoretical expectations, but it would be desirable to have substantially more data. Indeed, while it is possible to conduct a meta-analysis even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e highly uncertain estimates and CI often spanned zero). 
For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p-value = 0.049). Note that the CIs did not cross zero for seedling and sapling stage, despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistent effect in the coefficient reported for each stage across studies. 
As expected, for seed mass the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig. \ref{fig:fig3}c, CIs overlapping zero). However, the correlations between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). We could not distinguish between these alternatives. But the number of coefficients reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficients reported rather than to a much smaller numbers of replicates.

For maximum height and Aarea, correlations with relative growth rate were also consistent with expectations, but our ability to test all hypotheses was at times limited due to low or no sample size. For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig. \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.
For Aarea, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages, as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

>>>>

\bibliographystyle{mee}
\mstype{Research Article}
\keywords{growth strategy, life-history traits, maximum height, ontogenetic stage, photosynthetic rate, plant development, seed mass, specific leaf area, wood density}

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate is seen as a central element of its ecological strategy, and as determined by its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and growth, casting doubt on the capacity of some prominent traits to influence growth rate.

@@

\item {\it Synthesis}: We show for some traits, how trait-growth correlations change in a predictable way with plant size. Our understanding of plant strategies should shift away from describing species as having a fixed growth strategy throughout their life (on a continuous axis from slow to fast growth), in favour of a size-dependent growth trajectories.
\end{enumerate}

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients, and light for growth, and then ultimately replacing themselves with offspring. The idea that strategy differences among species might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who observed a wide spread of relative growth rates among seedlings growing under favourable conditions. Seedling potential relative growth rate has continued to be used as a major strategy indicator within Grime's CSR (Competitor -- Stress tolerator -- Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling relative growth rate into different components showed that specific leaf area (SLA) was typically the strongest source of variation between species \citep{Poorter:1989tx, Rees:2010gk}. During the 1990's a new style of ``trait ecology'' arose whereby measurable traits, including SLA, were used directly as strategy axes \citep{Westoby:2002ft}. This approach, using measurable traits, made broad comparisons possible across continents, latitudes, and thousands of species. Since then, functional traits have been seen as constructive tools to understanding plant growth strategies \citep{Westoby:2002ft}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of trait variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}.

Yet, recently a number of empirical studies indicate the correlation between traits and growth is not consistent. Consider SLA as an example. By definition, species with high SLA can deploy more leaf area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In seedlings growing under favourable conditions, SLA has repeatedly been found to be a strong predictor of relative growth rate, regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

@@

In this paper, we consider five prominent traits (specific leaf area = SLA, wood or stem tissue density, maximum leaf photosynthetic capacity per unit area = $A_{area}$, maximum or asymptotic height, and seed mass) and we investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions derived from a recently developed mechanistic model \citep{Falster:2011ii} (see ``Theoretical expectations'')?
\end{enumerate}
To address these questions we used a meta-analytic approach, compiling data from studies reporting correlation coefficients between traits and growth. While most studies investigating the link between traits and growth have not primarily been concerned with the effects of plant size, they have generally reported trait-growth correlations for identifiable ontogenetic stages (i.e. seedling, sapling, and adult). We therefore used ontogenetic stage as a surrogate for plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.
Using a meta-analysis approach, we were able to assess our hypotheses across a much broader number of species and vegetation types than would have been possible in an experimental study. We anticipated a sizeable literature on traits and growth to be available but we were uncertain about the representation of all trait-stage combinations. An additional goal of our study was therefore to identify gaps in the knowledge needed to improve our understanding of the link between functional traits and plant growth.

@@

The rate at which live biomass $B$ is produced by a plant is given by
\begin{equation}\label{eq:eq0}
\underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}} = \underbrace{\alpha_{\rm y}}_{\text{yield}}  \big(\underbrace{A_{\rm l} p}_{\text{photosynthesis}} - \underbrace{\sum_{i = {\rm l,s,r}} M_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i = \rm{l,s,r}} M_{i} k_{i}}_{\text{turnover}} \, ,
\end{equation}
where $\alpha_{\rm y}$ is the ratio of carbon fixed in mass per carbon assimilated, and $p$ is the assimilation rate of CO$_{2}$ per leaf area. Here, and in the following equations, $M_i$ denotes the live mass on the plant, $r_i$ the mass-specific respiration rate, and $k_i$ the turnover rate of different plant tissues; the subscripts $i = \rm{l,s,r,a,f}$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.)  Similarly $A_i$ denotes areas, of leaves ($i = l$) and of cross-sections of total stem and sapwood ($i =\rm{st,ss}$) respectively. Note that the live biomass $B = M_{\rm a} + M_{\rm f}$.  Applying the chain rule of differential calculus, the growth rate in plant mass, ($M_{\rm a} $), can be decomposed as
\begin{equation}\label{eq:eq1}
\underbrace{{\frac{\ud M_{\rm a} }{\ud t}}}_{\text{growth rate in mass}} = \underbrace{\frac{\ud M_{\rm a} }{\ud B}}_{\text{allocation to vegetative growth}}  \times \underbrace{\frac{\ud B}{\ud t}}_{\text{net live biomass production}},
\end{equation}
where $\frac{\ud M_{\rm a} }{\ud B}$ is the fraction of net live biomass increase that goes to vegetative growth rather than to production of reproductive tissues.

The growth rate in height can likewise be decomposed as the product of several meaningful terms:
\begin{equation}\label{eq:eq2}
\underbrace{\frac{\ud H}{\ud t}}_{\text{growth rate in height}} = \underbrace{\frac{\ud H}{\ud A_{\rm l}}}_{\text{architecture layout}} \times \underbrace{{\frac{\ud A_{\rm l}}{\ud M_{\rm a} }}}_{\text{leaf deployment per mass}} \times {\frac{\ud M_{\rm a} }{\ud B}} \times {\frac{\ud B}{\ud t}}
\end{equation}
where $\ud H / \ud A_{\rm l}$ is the growth in plant height
per unit growth in total leaf area. This term expresses the architectural
strategy of the plant. We assume it is unrelated to the traits considered in the current paper, in other words that it does not automatically change in response to changes in the traits. The term $\ud A_{\rm l} / \ud M_{\rm a} $
accounts for the marginal cost of deploying an additional unit of leaf
area, including construction of the leaf itself and various support
structures.  $\ud A_{\rm l} / \ud M_{\rm a} $
can itself be expressed as a sum of costs per unit leaf
area for constructing various support tissues,
\begin{equation}\label{eq:daldmt}
\frac{\ud A_{\rm l}}{\ud M_{\rm a} }
= \left(\frac{\ud M_{\rm l}}{\ud A_{\rm l}}
 +  \frac{\ud M_{\rm s}}{\ud A_{\rm l}} + \frac{\ud M_{\rm r}}{\ud A_{\rm l}}\right)^{-1}.
\end{equation}

Similarly growth rate in basal area can be expressed as the sum of growth in sapwood and heartwood, with the former term again decomposed as a product:
\begin{equation}\label{eq:eq3}
\begin{array}{ccccc}
\underbrace{{\frac{\ud A_{\rm st}}{\ud t}}}_{\text{basal area growth rate}} &=& \frac{\ud A_{\rm ss}}{\ud t} &+& \frac{\ud A_{\rm sh}}{\ud t} \\
&=& \underbrace{\frac{\ud A_{\rm ss}}{\ud A_{\rm l}}  \times {\frac{\ud A_{\rm l}}{\ud M_{\rm a} }} \times  {\frac{\ud M_{\rm a} }{\ud B}} \times {\frac{\ud B}{\ud t}}}_{\text{increase in total sapwood}} &+&  \underbrace{k_{\rm ss} \times \frac{\ud A_{\rm ss}}{\ud t}}_{\text{sapwood turnover}}.\\
\end{array}
\end{equation}
Here $\frac{\ud A_{\rm ss}}{\ud A_{\rm l}} $ is the  sapwood area per leaf area. The growth rate in diameter is then obtained via the geometric relationship between stem diameter and stem area,
\begin{equation}\label{eq:eq4}
\underbrace{\frac{\ud D}{\ud t}}_{\text{growth rate in diameter}} = \left(\pi \, A_{\rm st}\right)^{-0.5} \times {\frac{\ud A_{\rm st}}{\ud t}}.
\end{equation}

@@

The above equations are general, i.e. they do not depend on detailed assumptions about plant construction. To make qualitative predictions with respect to traits such as specific leaf area and wood density we require one additional assumption: that the mass of sapwood tissue ($M_{\rm ss}$) increases disproportionately with respect to the plant's leaf area ($A_{\rm l}$). This assumption is included in nearly every model of plant growth that has been developed, on the basis that as stems grow longer the total mass of sapwood associated with each leaf must be increasing \citep[e.g.][]{King-1999,Enquist:2007ek,Falster:2011ii}. It has the effect of changing the relative importance of different components within equations for $\frac{\ud B}{\ud t}$ (net live biomass production) and $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ (marginal cost of deploying additional leaf area) as a plant increases in size. Specifically, in eq. \ref{eq:eq0}, the costs of stem respiration and turnover increase faster than leaf photosynthesis, respiration or turnover, as the plant grows. Similarly, in eq. \ref{eq:daldmt}, the costs of constructing sapwood to support a given unit of leaf area, $\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$, increases faster than the cost of producing a unit of leaf area, $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ which is simply the inverse of specific leaf area. Whether plant growth rate increases with a specific trait, therefore depends on how the trait affects the different elements of \ref{eq:eq0}-\ref{eq:eq4} and how those effects change with size (Table \ref{tab:trade-offs}).

@@

\subsection*{H5: The relationship between $A_{area}$ and growth rate is expected to be positive across all plant sizes, particularly under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table \ref{tab:trade-offs}), at least at high light where $A_{area}$ is typically measured. This effect occurs by increasing  $\frac{\ud B}{\ud t} $ in eqs. (\ref{eq:eq1}-\ref{eq:eq3}). The positive effect of increasing $A_{area}$ (higher assimilation) may be offset by the costs (higher respiration), however, assuming both increase in proportion, and because assimilation is typically higher than respiration in high light, the net effect will be positive (Table \ref{tab:trade-offs}). It is uncertain how much this effect might translate to faster growth in low light.

@@

A big challenge when searching literature databases is to identify suitable studies from the vast corpus available. Initially we were interested to know whether data existed for a variety of traits; however the number of potential search terms made systemic searching for all these terms in Google Scholar impractical. A preliminary investigation of search results for a variety of traits on the Web of Science platform suggested that searching only on SLA and wood density -- two of the most widely measured traits -- would capture $\sim$80\% of search records that would be returned by searching for a wider range of traits (see Appendix \ref{app:literature-snapshot} in Supporting Information). We therefore narrowed our searching to only cover these two main traits.

@@

The 19 traits recorded (Fig. \ref{fig:fig1}) capture different components of plant ecological strategies. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and $A_{area}$ are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass and maximum height capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here wood density and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf versus other tissue types \citep{Lambers:1992bj}.

@@

We used categorical groupings by ontogenetic ``stage'' to test our hypotheses regarding the effects of traits on growth rates. The authors of the 103 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. Because these categories were established according to the expertise of the authors, we were concerned about the risk of assigning plants of similar size to different categories when compiling data across multiple studies and vegetation types. This risk turned out to be small: overall, we found that the original thresholds in sizeor age used by the authors to distinguish different stages were relatively compatible with thresholds we defined ourselves (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in stem diameter (at 10 cm). Comparing these values to the bivariate relationships observed between variables (age, height, diameter) in the ``Biomass And Allometry Database'' \citep{Falster:2015}, demonstrated that the thresholds identified above for each stage transition also aligned relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries.

@@

\subsection*{Representation of ontogenetic stages among 19 traits}
In total, correlations had been reported for at least two different stages across 17 traits out of the 19 targeted (Fig. \ref{fig:fig1}). We had thought it likely that for some traits, correlations would be reported only for one growth stage. But this limitation applied only for the net (mass) assimilation rates ($NAR_{mass}$) and vessel density, where studies appeared confined to seedling and adult stages respectively. However, for a majority of traits the number of observations across stage was highly unbalanced; for one-third of the 19 traits $>$ 75\% of correlations with growth reported come from the same stage (Fig. \ref{fig:fig1}).

@@

For wood density and seed mass, the available results are consistent with our theoretical expectations, but it would be desirable to have substantially more data. Indeed, while it is possible to conduct a meta-analysis even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e highly uncertain estimates and CI often spanned zero). 
For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p-value = 0.049). Note that the CIs did not cross zero for seedling and sapling stage, despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation or a good consistent effect in the coefficient reported for each stage across studies. 
As expected, for seed mass the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig. \ref{fig:fig3}c, CIs overlapping zero). However, the correlations between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). We could not distinguish between these alternatives. But the number of coefficients reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficients reported rather than to a much smaller numbers of replicates.

For maximum height and $A_{area}$, correlations with relative growth rate were also consistent with expectations, but our ability to test all hypotheses was at times limited due to low or no sample size. For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig. \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.
For $A_{area}$, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages, as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).
